Modern Statistical Methods --- Example Sheet 1

### 2020-10-21 distribution_theory/C1.tex
updatesssss

A special case of the previous example is the case $M = 0$: in this case the distribution simply becomes
\[
\ang{T_f, \phi} = \int_X f \phi \dd{x}. 
\]
Henceforth we will abuse notation: if $f \in \locint(X)$, then we will write $f$ instead of $T_f$, i.e., $\ang{f, \phi} = \int_X f\phi \dd{x}$.

@@

\subsubsection{Reflection and translation} 
For $\phi \in \DD(\RR^n), h \in \RR^n$, define the \emph{translation of $\phi$ by $h$} by 
\[
(T_h\phi)(x) \ceq \phi(x - h),
\]
and the \emph{reflection of $\phi$} by $\refl\phi(x) \ceq \phi(-x)$.

Extending the definitions of translation and reflection by duality yields the following:
\begin{definition}
    For $u \in \DD'(\RR^n)$, $h \in \RR^n$, $\phi \in \DD(\RR^n)$, define
    \[
    \ang{T_hu, \phi} \ceq \ang{u, T_{-h}\phi} \quad\text{and}\quad \ang{\refl u, \phi} \ceq \ang{u, \refl{\phi}}. 
    \]
\end{definition}

\begin{lemma}
    For $u \in \DD'(\RR^n)$, define $V_h \in \DD'(\RR^n)$ for $0 \neq h \in \RR^n$ by
    \[
    V_h \ceq \frac{T_{-h}u - u}{\norm{h}}
    \]
    If $(h_j) \subseteq \RR^n$ is a sequence for which $\lim_{j \to \infty} \frac{h_j}{\norm{h_j}} = m \in S^{n-1}$, then $V_{h_j} \to \sum_i m_i \pdv{u}{x_i}$ in $\DD'(\RR^n)$. 
\end{lemma}

\begin{proof}
    By definition, we can write $\ang{V_h, \phi} = \frac{1}{\norm{h}} \ang{u, T_h\phi - \phi}$. Now Taylor's theorem tells us that
    \[
    (T_h\phi - \phi)(x) = \phi(x - h) - \phi(x) = - \sum_i h_i \pdv{\phi}{x_i}(x) + R(x, h),
    \]
    where $R(x, h) = o(\norm{h})$ in $D(\RR^n)$ (exercise sheet 1, question 2).

@@

\subsubsection{Convolution}
For $\phi \in \DD(\RR^n)$, note that $(T_x \refl\phi)(y) = \refl\phi(y-x) = \phi(x-y)$. 
\begin{definition}
    For $u \in C^\infty(\RR^n), \phi \in \DD(\RR^n)$, we define the \emph{convolution} $u * \phi \colon \RR^n \to \CC$ as
    \[
    (u * \phi)(x) \ceq \int_{\RR^n} u(x-y) \phi(y) \dd{y} = \int_{\RR^n} u(y) \phi(x - y) \dd{y} = \ang{u, T_x \refl\phi}. 
    \]
    Since the RHS makes sense for any $u \in \DD'(\RR^n)$, we extend the definition this way: for $u \in \DD'(\RR^n), \phi \in \DD(\RR^n)$, we define the convolution $u * \phi$ as
    \[
    (u * \phi)(x) \ceq \ang{u, T_x\refl\phi}. 
    \]
\end{definition}

@@

>>>>

A special case of the previous example is the case $M = 0$: in this case the distribution simply becomes
\[
\ang{\tau_f, \phi} = \int_X f \phi \dd{x}. 
\]
Henceforth we will abuse notation: if $f \in \locint(X)$, then we will write $f$ instead of $\tau_f$, i.e., $\ang{f, \phi} = \int_X f\phi \dd{x}$.

@@

\subsubsection{Reflection and translation} 
For $\phi \in \DD(\RR^n), h \in \RR^n$, define the \emph{translation of $\phi$ by $h$} by 
\[
(\tau_h\phi)(x) \ceq \phi(x - h),
\]
and the \emph{reflection of $\phi$} by $\refl\phi(x) \ceq \phi(-x)$.

Extending the definitions of translation and reflection by duality yields the following:
\begin{definition}
    For $u \in \DD'(\RR^n)$, $h \in \RR^n$, $\phi \in \DD(\RR^n)$, define
    \[
    \ang{\tau_hu, \phi} \ceq \ang{u, \tau_{-h}\phi} \quad\text{and}\quad \ang{\refl u, \phi} \ceq \ang{u, \refl{\phi}}. 
    \]
\end{definition}

\begin{lemma}
    For $u \in \DD'(\RR^n)$, define $V_h \in \DD'(\RR^n)$ for $0 \neq h \in \RR^n$ by
    \[
    V_h \ceq \frac{\tau_{-h}u - u}{\norm{h}}
    \]
    If $(h_j) \subseteq \RR^n$ is a sequence for which $\lim_{j \to \infty} \frac{h_j}{\norm{h_j}} = m \in S^{n-1}$, then $V_{h_j} \to \sum_i m_i \pdv{u}{x_i}$ in $\DD'(\RR^n)$. 
\end{lemma}

\begin{proof}
    By definition, we can write $\ang{V_h, \phi} = \frac{1}{\norm{h}} \ang{u, \tau_h\phi - \phi}$. Now Taylor's theorem tells us that
    \[
    (\tau_h\phi - \phi)(x) = \phi(x - h) - \phi(x) = - \sum_i h_i \pdv{\phi}{x_i}(x) + R(x, h),
    \]
    where $R(x, h) = o(\norm{h})$ in $D(\RR^n)$ (exercise sheet 1, question 2).

@@

\subsubsection{Convolution}
For $\phi \in \DD(\RR^n)$, note that $(\tau_x \refl\phi)(y) = \refl\phi(y-x) = \phi(x-y)$. 
\begin{definition}
    For $u \in C^\infty(\RR^n), \phi \in \DD(\RR^n)$, we define the \emph{convolution} $u * \phi \colon \RR^n \to \CC$ as
    \[
    (u * \phi)(x) \ceq \int_{\RR^n} u(x-y) \phi(y) \dd{y} = \int_{\RR^n} u(y) \phi(x - y) \dd{y} = \ang{u, \tau_x \refl\phi}. 
    \]
    Since the RHS makes sense for any $u \in \DD'(\RR^n)$, we extend the definition this way: for $u \in \DD'(\RR^n), \phi \in \DD(\RR^n)$, we define the convolution $u * \phi$ as
    \[
    (u * \phi)(x) \ceq \ang{u, \tau_x\refl\phi}. 
    \]
\end{definition}

@@

Due to the previous corollary, we often call $u * \phi$ a \emph{regularisation} of $u$.

\begin{convention}
	If $u \in \DD'(X)$ and $\phi \in \DD(X)$, then instead of $\ang{u, \phi}$ we also write $\ang{u(t), \phi(t)}$ (or with any other dummy variable) when the variable used for $\phi$ is not directly clear. 
\end{convention}

\subsection{Density of $\DD(\RR^n)$ in $\DD'(\RR^n)$}
\begin{lemma}
	If $u \in \DD'(\RR^n)$ and $\phi, \psi \in \DD(\RR^n)$, then
	\[
	(u * \phi) * \psi = u * (\phi * \psi). 
	\]
\end{lemma}

\begin{proof}
	Fix $x \in \RR^n$. Now we write
	\begin{align*}
		((u * \phi) * \psi)(x) &= \int_{\RR^n} \ang{u(z), (\tau_y \refl\phi)(z)} \cdot \psi(x-y) \dd{y} \\
		&= \int_{\RR^n} \ang{u(z), (\tau_{x-y}\refl\phi)(z)} \cdot \psi(y) \dd{y} \\
		&= \int_{\RR^n} \ang{u(z), \psi(y) (\tau_{x-y} \refl\phi)(z)} \dd{y}.
	\end{align*}
We would like to interchange integral and application of $u$, and we will have to justify this using Riemann sums: 
\begin{align*}
\int_{\RR^n} \ang{u(z), \psi(y) (\tau_{x-y} \refl\phi)(z)} \dd{y} &= \lim_{\eps \downarrow 0} \sum_{m \in \ZZ^n} \ang{u(z), \psi(\eps m) \phi(x - z - \eps m)} \eps^n \\
&\overset*= \lim_{\eps\downarrow0} \ang{u(z), \sum_{m \in \ZZ^n} \psi(\eps m) \phi(x - z - \eps m)\eps^n } \\
&\overset{**}= \ang*{u(z), \int_{\RR^n} \psi(y) \phi(x - z - y) \dd{y}} \\
&= \ang{u(z), (\phi * \psi)(x - z)} = \ang{u(z), (\tau_x\widecheck{\phi * \psi})(z)} = (u * (\phi * \psi))(x). 
\end{align*}
Here, $*$ is by the fact that the sum is finite since $\psi$ has compact support, while $**$ is by sequential continuity of $u$ and the fact that the Riemann sum converges to the convolution integral \emph{in the space of test functions} (non-examinable fact). 
\end{proof}

\begin{theorem}
	If $u \in \DD'(\RR^n)$, there exists a sequence $(\phi_k) \subseteq \DD(\RR^n)$ such that $\phi_k \to u$ in $\DD'(\RR^n)$. 
\end{theorem}

\begin{proof}
	Fix $\psi \in \DD(\RR^n)$ with $\int_{\RR^n} \psi \dd{x} = 1$, and set $\psi_k(x) \ceq k^n \psi(kx)$. Note that $\int_{\RR^n} \psi_k \dd{x} = 1$.

	Now, fix any $\chi \in \DD(\RR^n)$ with $\chi \equiv 1$ on $\qty{\norm{x} < 1}$ and $\chi \equiv 0$ on $\qty{\norm{x} < 2}$. Define $\chi_k(x) \ceq \chi(x/k)$, so that $\lim_{k\to\infty} \chi_k(x) = 1$ for all $x$. We will set
	\[
	\phi_k(x) \ceq \chi_k(x) (u * \psi_k)(x). 
	\]
	Clearly we have $\phi_k \in \DD(\RR^n)$ since each $\chi_k$ has compact support.

	Note that $\ang{u, \phi} = (u * \refl\phi)(0)$ by definition. Now, take any $\theta \in \DD(\RR^n)$, then we have

	\begin{align*}
		\ang{\phi_k, \theta} &= \ang{\chi_k (u * \psi_k), \theta} = \ang{u * \psi_k, \chi_k \theta} = \qty[ (u * \psi_k) * \refl{\chi_k\theta}](0) \\
		&= \qty[u * \qty(\psi_k * \refl{\chi_k \theta})](0).
	\end{align*}

	Now we compute $\psi_k * \refl{\chi_k\theta}$: note that 
	\begin{align*}
		(\psi_k * \refl{\chi_k\theta})(x) &= \int_{\RR^n} k^n\psi(k(x-y)) \chi\qty(-\frac yk) \theta(-y) \dd{y} \\
		&= \int_{\RR^n} \psi(y) \chi\qty(\frac{y}{k^2} - \frac xk) \theta(\frac yk - x) \dd{y} \\
		&= \theta(-x) + R_k(-x) = (\refl{\theta + R_k})(x)
	\end{align*}
where $R_k(x) = \int_{\RR^n} \psi(y) \qty[ \chi\qty(\frac y{k^2} + \frac xk) \theta\qty(\frac yk + x )- \theta(x)] \dd{y}$.

So \[
\ang{\phi_k, \theta} = (u * (\refl{\theta + R_k}))(0) = (u *\refl\theta)(0) + (u * \refl{R_k})(0) = \ang{u, \theta} + \ang{u, R_k}. 
\]

We must now only prove that $R_k \to 0$ in $\DD(\RR^n)$, and then by sequential continuity it follows that $\phi_k \to u$  in $\DD'(\RR^n)$. 
\end{proof}

### 2020-10-21 inverse_problems/inverse_ES1.tex
updatesssss

	% \item From Fourier theory it is known that $\qty{s_n(x) = \sqrt 2\sin(nx) : n \in \NN}$ and $\qty{1} \cup \qty{c_n(x) = \sqrt 2\cos(nx) \mid n \in \NN}$ are orthonormal bases of $L^2[0, 1]$. 
	%Note that 
	%\[
	%(A s_n)(y) = \int_0^y s_n(x) \dd{x} = \sqrt 2 \int_0^y \sin(nx) \dd{x} = \frac{\sqrt 2}{n} \qty(1 - \cos(ny)) = \frac{\sqrt 2}{n} c_0(y) - \frac1n c_n(y) 
	%\]
	\item Clearly, any operator in $\Ran(A)$ is continuous by the fundamental theorem of calculus. Since $f$ is not continuous (also not up to a measure zero set), we conclude $f \notin \Ran(A)$ and therefore $f \in \clos{\Ran(A)} \setminus \Ran(A)$.

	\end{enumerate}
\end{solution}
\end{document}

>>>>

	% \item From Fourier theory it is known that $\qty{s_n(x) = \sqrt 2\sin(nx) : n \in \NN}$ and $\qty{1} \cup \qty{c_n(x) = \sqrt 2\cos(nx) \mid n \in \NN}$ are orthonormal bases of $L^2[0, 1]$. 
	%Note that 
	%\[
	%(A s_n)(y) = \int_0^y s_n(x) \dd{x} = \sqrt 2 \int_0^y \sin(nx) \dd{x} = \frac{\sqrt 2}{n} \qty(1 - \cos(ny)) = \frac{\sqrt 2}{n} c_0(y) - \frac1n c_n(y) 
	%\]
	\item To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$.

	Note that 
	\begin{align*}
		\ang{Au, v} &= \int_0^1 (Au)(y)\cdot v(y) \dd{y} \\
		&= \int_0^1 \int_0^y u(x) \dd{x} v(y) \dd{y} \\
		&= \int_0^1 \int_0^y u(x) v(y) \dd{x} \dd{y} \\
		&= \int_0^1 \int_x^1 u(x) v(y) \dd{y} \dd{x} \\ 
		&= \int_0^1 u(x) \int_x^1 v(y) \dd{y} \dd{x} \\
		&= \ang{u, A^* v}
	\end{align*}
where $v(x) = \int_x^1 v(y) \dd{y}$.  Therefore, we find
\begin{align*}
	(AA^* u)(x) = \int_0^y \int_x^1 u(x) \dd{x} \dd{y}
\end{align*}
	\end{enumerate}
\end{solution}
\end{document}

### 2020-10-21 modern_statistical_methods/C1.tex
updatesssss

This gives our first inequality $MSPE \leq \frac{\sigma^2}{n} \sum_{i=1}^n \frac{d_i^2}{(d_i + \lambda)^2} + \frac\lambda{4n}$. For the second inequality, note that $\frac{d_i^2}{(d_i + \lambda)^2} \leq 1$ and also $\frac{d_i^2}{(d_i + \lambda)^2} \leq \frac{d_i^2}{4 d_i\lambda} = \frac{d_i}{4\lambda}$. Therefore, we can write $\frac{d_i^2}{(d_i + \lambda)^2} \leq \frac1\lambda \min(\lambda, \frac{d_i}{4})$ which proves the second inequality. 
\end{proof}

>>>>

This gives our first inequality $\MSPE \leq \frac{\sigma^2}{n} \sum_{i=1}^n \frac{d_i^2}{(d_i + \lambda)^2} + \frac\lambda{4n}$. For the second inequality, note that $\frac{d_i^2}{(d_i + \lambda)^2} \leq 1$ and also $\frac{d_i^2}{(d_i + \lambda)^2} \leq \frac{d_i^2}{4 d_i\lambda} = \frac{d_i}{4\lambda}$. Therefore, we can write $\frac{d_i^2}{(d_i + \lambda)^2} \leq \frac1\lambda \min(\lambda, \frac{d_i}{4})$ which proves the second inequality. 
\end{proof}

We can now ask ourselves if kernel ridge regression is optimal in any sense. To this end, we define $\hat \mu_i \ceq d_i / n$ and $\lambda_n \ceq \lambda/n$. Then we can rewrite the upper bound from our precious theorem as

\[
\MSPE \leq \frac{\sigma^2}{n \lambda_n} \sum_{i=1}^n \min(\frac{\hat \mu_i}{4}, \lambda_n) + \frac{\lambda_n}{4} \eqqcolon \delta_n(\lambda_n).  
\]

Now, instead of taking $x_1, \dotsc, x_n \in \XX$ to be fixed, assume that they are i.i.d.\ random variables (then $K$ is random, so the $\hat\mu_i$ are random as well), and taking an expectation on both sides yields
\[
\EE[\MSPE] \leq \EE[\delta_n(\lambda_n)]. 
\]

We want to bound $\delta_n(\lambda_n)$ by some function of $n$. For this we will use Mercer's theorem:
\begin{theorem}[Mercer]
	Let $\XX = [a, b]$. If $k$ is a continuous kernel on $\XX$, then there exists an orthonormal basis $(e_i)$ of $L^2[a, b]$ such that
	\[
	k(x, x') = \sum_{j=1}^\infty \mu_j e_j(x) e_j(x')
	\]
	as well as a sequence of nonnegative eigenvalues $(\mu_j)$ such that
	\[
	\int_a^b k(x, x') e_j(x) \dd{x} = \mu_j e_j(x). 
	\]
\end{theorem}

Applying Mercer's theorem to $k$, it can be proved that for some $C> 0$ we have
\[
\EE\qty[\frac1n \sum_{i=1}^n \min(\frac{\hat \mu_i}{4}, \lambda_n)] \leq \frac Cn \sum_{i=1}^\infty \min(\frac{\mu_i}{4}, \lambda_n).
\]

\begin{example}
	Let $k$ be the Sobolev kernel. Then the eigenvalues satisfy 
	\[
	\frac{\mu_j}{4} = \frac{1}{\pi^2(2j - 1)^2}, 
	\]
	and with some calculations it can be shown that $\sum_{i=1}^\infty \min(\frac{\mu_i}{4}, \lambda_n) = O(\sqrt{\lambda_n})$ as $\lambda_n \to 0$. Combining this with the rest of the bound shows
	\[
	\EE[\delta_n(\lambda_n)] = O\qty(\frac{\sigma^2}{n} \lambda_n^{-1/2} + \lambda_n). 
	\]
	The optimal scaling $\lambda_n \sim (\sigma^2/n)^{2/3}$ gives an error rate of order $(\sigma^2/n)^{2/3}$.

	It has been shown that this is in fact the optimal rate for estimating a function $f^0 \in \HH$ (up to multiplicative constants). 
\end{example}

\subsection{Other kernel machines}

\subsubsection{The support vector machine}
Consider a \emph{classification problem} with data $x_1, \dotsc, x_n \in \RR^p$ and binary response $Y_i \in \qty{\pm 1}$. Furthermore assume that $x_1, \dotsc, x_n$ are separated by a hyperplane through the origin, that is, for some  $\beta \in \RR^p$ we have $Y_i x_i\T \beta > 0$ for all $i$.

To choose between different planes that separate the classes, we pick the hyperplane with the highest margin: that is, we compute
\[
\max_{\beta \in \RR^p, M \geq 0} M \qquad\text{such that } \frac{Y_i x_i\T\beta}{\norm{\beta}} \geq M \quad (i = 1,\dotsc, n). 
\]
When the classes are not separable, we can penalise according to the distance of a point ``over the margin''. The penalty should be 0 if $x$ is on the correct side and should equal the distance over the boundary otherwise, measured in units of $M$. In this case, by considering minimisation of $1/M^2$ instead of maximisation of $M$ we find
\[
\argmin_{M \geq 0, \beta \in \RR^p} \frac1{M^2} + \lambda \sum_{i=1}^n \qty(1 - \frac{Y_i x_i\T \beta}{\norm{\beta} M})_+. 
\]
Since the above equation is independent of the norm of $\beta$, we may set $\norm{\beta} = 1/M$ and remove $M$ from the equation entirely, thus obtaining
\[
\argmin_{\beta \in \RR^p} \norm{\beta}^2 + \lambda \sum_{i=1}^n (1 - Y_i x_i\T \beta)_+.
\]
Replacing $\lambda$ by $\frac1\lambda$ and multiplying the equation with $\lambda$ we obtain
\[
\argmin_{\beta \in \RR^p} \sum_{i=1}^n (1 - Y_i x_i\T \beta)_+ + \lambda \norm{\beta}^2. 
\]
Finally, if we lift the restriction that the hyperplane must go through the origin, we find the objective function
\begin{equation} \label{eq:support_vector_classifier}
(\hat \mu, \hat\beta) = \argmin_{\mu \in \RR, \beta \in \RR^p} \sum_{i=1}^n (1 - Y_i (x_i\T\beta + \mu))_+ + \lambda  \norm{\beta}^2. 
\end{equation}
The solution to this problem is known as the \emph{support vector classifier}.

\paragraph{Introducing kernels}
Now we introduce kernels to the problem. Letting $k$ be the linear kernel and $\HH$ the corresponding RKHS, we can rewrite \cref{eq:support_vector_classifier} as
\[
(\hat \mu, \hat f) = \argmin_{\mu \in \RR, f \in \HH} \sum_{i=1}^n (1 - Y_i\T (f(x_i) + \mu))_+ + \lambda \norm{f}^2.
\]
Suppose we change $\HH$ to a different RKHS with a different kernel $k$. By a variant of the representer theorem (Example Sheet 1, question 10) the solution is equivalent to
\[
(\hat\mu, \hat\alpha) = \argmin_{\mu \in \RR, \alpha \in \RR^n} \sum_{i=1}^n (1 - Y_i (K_i\T \alpha + \mu))_+ + \lambda \alpha\T K\alpha, 
\]
and the solution to the above problem is called the \emph{support vector machine}.  Predictions at a new point $x^*$ are given by
\[
\sign(\hat\mu + \sum_{i=1}^n \hat\alpha_i k(x^*, x_i)) \in \pm 1. 
\]
\begin{remark}
Note that the support vector machine can correspond to a nonlinear boundary: the boundary is a hyperplane \emph{in the corresponding RKHS}, but can be nonlinear in the space $\XX$ where the data points live. This also makes it easy to overfit. 
\end{remark}
